### 2023-06-06 problem_2_dynamic.tex

\subsection{Постановка задачі}
Розглядається пружна така сама прямокутна область як і в попередній задачі