### 2023-07-09 pub_list.tex

\begin{enumerate}
    \item D. Nerukh, O. Pozhylenkov, N. Vaysfeld (2019) Mixed plain boundary value problem of elasticity for a rectangular domain.
    25-th International Conference Engineering Mechanics. 2019, May 13-16, Svratka, Czech Republic. p. 255
    \item O. V. Pozhylenkov (2019) The stress state of a rectangular elastic domain. Researches in Mathematics and Mechanics, Volume 24, Issue 2(34), pp. 88-96
    \item Пожиленков О. В. Вайсфельд Н. Д. (2019) Мішана крайова задача теорії пружності для прямокутної області. Математичні проблеми механіки неоднорідних структур, випуск 5, Львів, ст. 30-32
    \item O. Pozhyenkov, N. Vaysfeld (2020) Stress state of a rectangular domain with the mixed boundary conditions. Procedia Structural Integrity, Volume 28, pp. 458-463
    \item O. Pozhyenkov, N. Vaysfeld (2021) Stress state of an elastic rectangular domain under steady load. Procedia Structural Integrity, Volume 33, pp. 385-390
    \item O. Pozhylenkov, N. Vaysfeld (2022) Dynamic mixed problem of elasticity for a rectangular domain. Recent trends in Wave Mechanics and Vibrations, pp. 211-218
\end{enumerate}