### 2023-07-14 abstract.tex
update

\textit{Pozhylenkov O. V.}
Plane mixed problems of elasticity for a rectangular domain - Manuscript.

A dissertation submitted for the degree of Doctor of Philosophy in the field of 113 «Applied Mathematics» (11 - «Mathematics and Statistics») - Odessa I.I. Mechnikov National University, Odessa, 2023.

The plane mixed boundary value problems of theory elasticity were solved for an elastic rectangular domain subjected to static and dynamic loads.
By applying the integral finite sin-cos Fourier transformation, the original problem was reduced to a one-dimensional boundary value problem, which was reformulated in the form of a vector boundary value problem.
The solution to this problem was constructed as a superposition of the general solution of the homogeneous vector equation and the particular solution of the nonhomogeneous equation.
The solution of the homogeneous vector equation was obtained using matrix differential calculus and represented using the fundamental matrix solution system of the corresponding homogeneous matrix equation.
The particular solution of the nonhomogeneous vector equation was found using the Green's matrix function representation.
The application of inverse Fourier transformation and the realization of the separation of weakly convergent parts of the integral provide the field of displacements and stresses through an unknown function - the boundary value of displacements along the end of the rectangular region.
To find this function under the boundary condition, a singular integral equation was derived and solved using the method of orthogonal polynomials.
The stress state of the medium was investigated for various types of loading and different geometric dimensions of the rectangular region.

\textit{Key words:}
rectangular domain, dynamic problem, Fourier transformation, matrix Green's function, singular integral equation, method of orthogonal polynomials.